---
title: "Getting Started: OFH Synthetic Cohort Generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started: OFH Synthetic Cohort Generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

This vignette shows how to generate synthetic cohort datasets for method development before using real health data.

The package-style API supports:

- configurable cohort size
- reproducible generation via seed
- optional ICD-10 / OPCS4 / BNF code restrictions

## 1. Load the package

```{r load-api, eval = FALSE}
library(OFHSimulatedData)
```

## 2. Generate a basic cohort

```{r basic-run, eval = FALSE}
out <- generate_ofh_cohort(
  n = 1000,
  seed = 123
)

names(out)
```

This returns a named list of data frames and writes CSVs to an output folder in your current working directory.

If you run this interactively, the generated data frames are also available in your R environment (for example `questionnaire_data`, `clinic_measurements_data`, `nhse_inpat_data`).

## 3. Restrict to specific code lists

```{r code-lists, eval = FALSE}
out <- generate_ofh_cohort(
  n = 1000,
  seed = 123,
  icd10 = c(
    I210 = "STEMI of anterolateral wall",
    I500 = "Congestive heart failure"
  ),
  opcs4 = c(
    K401 = "Percutaneous transluminal balloon angioplasty of coronary artery"
  ),
  bnf_codes = c("0212000B0", "0601023A0")
)
```

You can also provide code files (TXT or CSV) with one code per line:

```{r code-files, eval = FALSE}
out <- generate_ofh_cohort(
  n = 1000,
  seed = 123,
  icd10_file = "icd10_codes.txt",
  opcs4_file = "opcs4_codes.csv",
  bnf_codes_file = "bnf_codes.txt"
)
```

## 4. Use the OOP interface directly

```{r oop-run, eval = FALSE}
sim <- OFHCohortSimulator$new(project_root = ".", seed = 123)

sim$set_code_pools(
  icd10 = c(I210 = "STEMI of anterolateral wall"),
  opcs4 = c(K401 = "Percutaneous transluminal balloon angioplasty of coronary artery"),
  bnf_codes = c("0212000B0", "0601023A0")
)

out <- sim$run_all(n = 800)
```

## 5. Practical tips for researchers

- Start with small `n` (for example, 200 to 1000) while developing.
- Fix `seed` for reproducibility during method testing.
- Check row counts and `pid` linkage assumptions in your analysis scripts.
- Expand code lists as your phenotype definitions evolve.

## 6. Notes

- Some datasets are intentional subsets of the full cohort.
- Questionnaire output includes a small v1 proportion by design.
- Primary care meds include prescribed-but-not-dispensed rows.
